\large{\bf Large effects and the infinitesimal model }

### 2023-12-05 review-responses.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript,
``Large effects and the infinitesimal model''.
In response to the reviewers' comments
we have made quite a few additions to the text
which we hope will make the paper more accessible to a wider audience.
We hope you agree!

@@

\begin{point}{}
More generally, I found it a bit hard to follow the overall logical flow of section 4.1, especially once the authors launch into an exploration of the ``reproduction'' and ``noise'' terms (pages 9 and 10).
    It would be useful to clarify what the biological/intuitive meaning of the noise and reproduction terms at the outset. It would also be useful to state (at the beginning) what the main goal of these explorations is: (is the goal to identify when both terms ``have well-posed limits independent of the other'' as stated at \llname{limit_remark}), and also to summarise (at the end) to what extent one can do this in all generality.
\end{point}

\reply{
    Good suggestion; we've added more clarifying points;
    see \llname{clarification1}
    and \llname{clarification2}
    (and \llname{clarification3}).
}

\begin{point}{}
    Last inequality above \revref: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution? If the former, then it'd be good to say this explicitly. If the latter, then this does not quite answer the question posed at the beginning of the section about how much information knowing the trait value of an individual gives about underlying allelic effects.
\end{point}

\reply{
    It is both --
    the calculation is, as the reviewer points out, about the largest of $M$ iid Cauchys,
    but those are the alleles carried by a parent in the initial generation.
    Furthermore, as explained at the beginning of this section \llname{crucial_step},
    this calculation about iid effects is indeed the key step
    (in \citet{fisher1918correlation} and \cite{barton2017infinitesimal})
    in describing the distribution of trait values in offspring.
    We've added an sentence that we think should remind the reader
    why we have in fact answered the question \revref.
}

@@

\begin{quote}
Anyway, the results
in Sect 4 are of interest because they strongly constrain the possibilities to justify such
‘Gaussian-population approximation’ by any infinitesimal-like model (essentially to
mixtures of Cauchy distributions. The noted relation to blending inheritance is intriguing!).
My main concern is that especially this section needs are clearer description of the model,
and clear distinction between this ‘maximalist’ model (which has little theoretical
support) and the model investigated by Barton et al (2017). Also a much clearer summary
and discussion of the implications of the results shown here are needed. It should not be
assumed that readers have read Turelli (2017) and even less that they have studied
\citet{barton2017infinitesimal} in considerable detail!
\end{quote}

Thanks for the suggestions.
We also thought the section was useful (and also intruiging!),
which is why we left the section in.
(Also note it motivates the form of selection in our simulations.)
We share the reviewer's concern about context,
and have added some additional text
while trying not to get into the weeds too much
-- in particular, see the new text around \llname{clarification3}.

\begin{quote}
In Section 5, it is pointed out why the Barton et al. (2017) fails in the case of ‘Cauchy
noise’ (and presumably for any other limit law with $\alpha < 2$). In the absence of giving the
estimate in App D in Barton et al, the explanation around \llname{approx_eqn} is difficult to
follow. Please provide their estimate for easier comparison (do not assume that readers
studied Barton et al; assume that they scanned (parts) of it).
\end{quote}

We deliberately avoided giving the analogous estimate from Barton et al. (2017), as the expression is complicated and far from transparent (its definition spans several paragraphs). As such, we instead chose to focus on its salient features, namely that scales as $\frac{1}{\sqrt{M}}$ and grows (approximately) proportionally to the magnitude of the trait.  We have added a reference to the relevant appendix in Barton et al. (2017) for those sufficiently motivated, and have attempted to give some more intuition for the consequences of the difference in the error bounds.

@@

\begin{point}{\revref}
This discussion is too vague. At least, adding a summary of the main findings
would be appropriate (always noting for which version of the infinitesimal model).
\end{point}

\reply{
    We have added a good bit more discussion, including a summary \revref.
}

\begin{point}{\revref}
Does `Lévy stable distribution’ refer to the family of $\alpha$-stable
distributions? (I don’t think it has not been used before.)
\end{point}

\reply{
    That's right; we've edited this bit, which no longer says ``Lévy''. \revref
}

### 2023-12-05 writeup.tex
final? edits

Much of the work to 
understand the evolution of quantitative traits in sexual species
has focused on models in quantitative genetics
that describe evolution of trait values without explicitly representing underlying genetic variation,
but in a manner deriving from our understanding of how genetic variation works in practice.
Such models are therefore
interestingly intermediate between mechanistic models of genetics and purely phenomenological.
% The action of natural selection on genetic variants in the absence of genetic variation is relatively well-understood (cite ewens),
% but tractable mathematical models of evolution for recombining species are elusive,
% at least with some results of when the model might be a good description of reality.
% For instance, a great deal of conceptual understanding and even quantitative predictions
% have been made assuming that the distribution of trait values in a population is Gaussian (cite Lande etcetera)
% despite the absence of a plausible model that would make this true (cite turelli and barton).
One concrete model of heritable trait evolution is the \emph{infinitesimal model},
which assumes a simple, explicit form for how the genetic component of offsprings' traits are determined by their parents:
they are Normally distributed, with the mean for each offspring equal to the average of the parents' values,
and a covariance independent of the parental trait values (but that depends on how they are related).
\citet{barton2017infinitesimal} give an overview of the complex history of the model, \revpoint{1}{12}
which is used in or adjacent to a good deal of theoretical work on quantitative genetics
\citep[e.g.,][]{fisher1918correlation,bulmer1971effect,lande1975maintenance}.
\citet{barton2017infinitesimal} focus on giving a formal proof of conditions
under which this would be a good approximation to reality,
which they extended to include dominance in \citet{barton2022infinitesimal}.

As suggested by its name,
under the infinitesimal model the effects of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical understanding of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many -- if not most -- lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps a trait is affected by mutations in a number of genomic regions,
and the distribution of effect sizes within each region is Normal, \revpoint{1}{13}
but with a standard deviation that varies with ``proximity'' of that region to the trait
(e.g., regions that are further ``upstream'' have higher standard deviations).
Such mixtures can easily have large proportions of extreme values.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
On the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{1}{11}

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, some important traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race
\citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus,
and $M$ is the number of loci. \revpoint{2}{2}
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{Mendelian sampling term}
(or occasionally \emph{Mendelian noise}), $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}
Note in particular that
the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|$.

@@

A possible concern is that tail exponents are difficult to estimate accurately, \revpoint{1}{4}
but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
show that our estimates are unbiased but only accurate to within roughly $\pm 0.4$.
So, the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
a direct estimate of the underlying distribution of $\alpha$ values,
but is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).

@@

We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$ as $M \to \infty$,
then with $\alpha \ge 2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
(The main result in \citet{barton2017infinitesimal} is that the trait values in generation $t$
are Gaussian conditioned on the traits in generation $t-1$,
but as discussed in that paper,
this implies that the traits of any set of individuals is multivariate Gaussian,
with covariance matrix determined by the pedigree.
Note, however, that this does \emph{not} imply that the distribution of trait values
across the population is Gaussian, as a collection of correlated Gaussians is not itself Gaussian.)

Perhaps surprisingly, there is a converse --
in at least some situations,
independence of the Mendelian sampling term \emph{implies} a Gaussian distribution.
To see this (and to see exactly what we mean by this),
divide the parental loci into groups according to the results of inheritance,
defining
\begin{align*}
    U_M^{[1]} &= \sum_{\ell=1}^M X_\ell \eta_\ell^{[1]}  ,
    \qquad \qquad
    V_M^{[1]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[1]} ,  \\
    U_M^{[2]} &= \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[2]} ,
    \qquad \qquad
    V_M^{[2]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[2]}  ,
\end{align*}
so that 
$Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, while
$Z_M = U_M^{[1]} + U_M^{[2]}$ is the transmitted portion of the parental genomes,
and we can define $V_M = V_M^{[1]} + V_M^{[2]}$ to be the untransmitted portion.
Furthermore,
\begin{align*}
    \bar Z_M = \frac{Z_M + V_M}{2}
    \qquad \text{ and } \qquad
    R_M = \frac{Z_M - V_M}{ 2 } .
\end{align*}

@@

The Kac-Bernstein theorem states that if two independent random variables
have the property that their sum and difference are independent of each other,
then the random variables have a joint Gaussian distribution
(\citet{kac1939characterization,bernstein1941property},
and see \citep{kagan1973characterization} for many much more general results).
Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
of $Z_M$ and $V_M$, and $Z_M$ and $V_M$ are independent,
if we assume that also $R_M$ and $\bar Z_M$ are independent,
then both are therefore jointly Gaussian.

@@

Note that under the conditions of the proposition, \llabel{pop_not_gaussian}
the parental traits, $Z^{[1]}_M$ and $Z^{[2]}_M$,
are also (jointly) Gaussian.
(This was not an assumption, however.)
If parents are chosen uniformly, then
the distribution of, say, $Z^{[1]}$ is just the population's trait distribution.
This seems wrong, since it is well-known that forces such as selection or migration
can make the population distribution non-Gaussian.
However, there is no contradiction:
we have stated the proposition only in terms of the \emph{first} generation
of offspring of unrelated parents \citep[as in][]{barton2017infinitesimal}.
Proposition~\ref{prop:parental_contribs} does not imply that
the distribution of traits in the offspring generation is Gaussian:
unless all offspring are unrelated,
it will be a collection of correlated Gaussians,
which is not itself Gaussian.
(As an extreme example, suppose that half the offspring are produced by the two parents
with largest trait values, and the other half by the parents with the smallest trait values;
then although each offspring's distribution about the midparent is Gaussian,
the offspring's population distribution is a mixture of two Gaussians
about these two midparents.)
What about further generations?
A similar statement to the proposition
holds for the offspring of two parents in an arbitrary pedigree,
using the Skitovich-Darmois theorem \citet{kagan1973characterization}
% except that the portion of the genome shared in common by the parents is unconstrained.
-- however, there is nothing to be gained by the additional generality.

@@

As observed in \cite{turelli2017commentary},
there are at least three interpretations of what is called ``the infinitesimal model''. 
The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation,
assumes parental trait-independent Gaussian Mendelian sampling terms,
and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
As also discussed in \citet{turelli2017commentary}, \revpoint{1}{8}
this assumption is rather fragile --
a Gaussian distribution for the population only emerges under weak or Gaussian selection.
% However, the model has been used in influential work.
In principle, another factor that could produce a non-Gaussian population distribution
is if the Mendelian sampling terms are not Gaussian, as we consider here --
so, it is interesting to look for other situations where the population distribution is analytically tractable.

@@

First, consider the case without selection, so $\mu$ is a constant
which we may take equal to 1. 
(To remind the reader, our goal is therefore to describe \llabel{clarification1}
what forms the population distribution might take in a neutral model
if we assume only that the Mendelian sampling terms are independent of parental values,
and nothing else about the distribution of these terms or the initial population distribution.) 
Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta) ,
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary function, $\varphi(\theta)$, under repeated applications of $F$,
\ie \revpoint{2}{7}
\begin{equation}\label{eq:fpeq}
\begin{aligned}
	\phi(\theta) 
    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
\end{aligned}
\end{equation}
% Iterating $F$ $n$ times, we get
% \[
% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=0}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
% \]
% so that
% \[
% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
% \]
is a fixed point whenever this limit exists, whereas any fixed point $\phi$ of $F$ satisfies  $\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\phi)(\theta)$.  As we are interested in identifying those fixed points that are characteristic functions, we henceforth assume, without loss of generality, that the $\varphi$ is itself a characteristic function (if $\phi$ exists then we can always take $\varphi = \phi$) and that the limit $\lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$ exists.

>>>>

Much of the work to 
understand the evolution of quantitative traits in sexual species
has focused on models in quantitative genetics
that describe evolution of trait values without explicitly representing underlying genetic variation,
but in a manner deriving from our understanding of how genetic variation works in practice.
Such models are therefore
interestingly intermediate between mechanistic models of genetics and purely phenomenological.
% The action of natural selection on genetic variants in the absence of genetic variation is relatively well-understood (cite ewens),
% but tractable mathematical models of evolution for recombining species are elusive,
% at least with some results of when the model might be a good description of reality.
% For instance, a great deal of conceptual understanding and even quantitative predictions
% have been made assuming that the distribution of trait values in a population is Gaussian (cite Lande etcetera)
% despite the absence of a plausible model that would make this true (cite turelli and barton).
One concrete model of heritable trait evolution is the \emph{infinitesimal model},
which assumes a simple, explicit form for how the genetic component of offsprings' traits are determined by their parents:
they are Normally distributed, with the mean for each offspring equal to the average of the parents' values,
and a covariance independent of the parental trait values (but that depends on how they are related).
\citet{barton2017infinitesimal} give an overview of the complex history of the model, \revpoint{1}{12}
which is used in or adjacent to a good deal of theoretical work on quantitative genetics
\citep[e.g.,][]{fisher1918correlation,bulmer1971effect,lande1975maintenance}.
\citet{barton2017infinitesimal} focus on giving a formal proof of conditions
under which this would be a good approximation to reality,
a proof they later extended to include dominance in \citet{barton2022infinitesimal}.

As suggested by its name,
under the infinitesimal model the effect of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical studies of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many -- if not most -- lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps a trait is affected by mutations in a number of genomic regions,
and the distribution of effect sizes within each region is Normal, \revpoint{1}{13}
but with a standard deviation specific to the region that varies with ``proximity'' of that region to the trait
(e.g., regions that are further ``upstream'' have higher standard deviations).
Such mixtures can easily have large proportions of extreme values.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
On the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
and so the sum of $n$ copies, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
(On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$.) \revpoint{1}{11}

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, some important traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
only appear heritable through non-genetic means
such as systematic inequalities in resource distribution partitioned by class or race
\citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus,
and $M$ is the number of loci. \revpoint{2}{2}
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{Mendelian sampling term}
(or occasionally \emph{Mendelian noise}), $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}
Note in particular that
the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|/2$.

@@

A possible concern is that tail exponents are difficult to estimate accurately, \revpoint{1}{4}
but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
show that our estimates are unbiased but only accurate to within roughly $\pm 0.4$.
So, the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
a direct estimate of the underlying distribution of $\alpha$ values,
but it is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).

@@

Now we return to the theory.
We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$ as $M \to \infty$,
then with $\alpha \ge 2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
(The main result in \citet{barton2017infinitesimal} is that the trait values in generation $t$
are Gaussian conditioned on the traits in generation $t-1$,
but as discussed in that paper,
this implies that the traits of any set of individuals is multivariate Gaussian,
with covariance matrix determined by the pedigree.
Note, however, that this does \emph{not} imply that the distribution of trait values
across the population is Gaussian, as a collection of correlated Gaussians is not itself Gaussian.)

Perhaps surprisingly, there is a converse --
in at least some situations,
independence of the Mendelian sampling term \emph{implies} a Gaussian distribution.
To see this (and to see exactly what we mean by this),
divide the parental loci into groups according to the results of inheritance,
defining
\begin{align*}
    U_M^{[1]} &= \sum_{\ell=1}^M X_\ell \eta_\ell^{[1]}  ,
    \qquad \qquad
    V_M^{[1]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[1]} ,  \\
    U_M^{[2]} &= \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[2]} ,
    \qquad \qquad
    V_M^{[2]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[2]}  ,
\end{align*}
so that the parental traits are
$Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, while
$Z_M = U_M^{[1]} + U_M^{[2]}$ is the transmitted portion of the parental genomes,
and we can define $V_M = V_M^{[1]} + V_M^{[2]}$ to be the untransmitted portion.
Furthermore,
\begin{align*}
    \bar Z_M = \frac{Z_M + V_M}{2}
    \qquad \text{ and } \qquad
    R_M = \frac{Z_M - V_M}{ 2 } .
\end{align*}

@@

The Kac-Bernstein theorem states that if two independent random variables
have the property that their sum and difference are independent of each other,
then the random variables have a joint Gaussian distribution
(\citet{kac1939characterization,bernstein1941property},
and see \citet{kagan1973characterization} for many much more general results).
Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
of $Z_M$ and $V_M$, and $Z_M$ and $V_M$ are independent,
if we assume that also $R_M$ and $\bar Z_M$ are independent,
then both are therefore jointly Gaussian.

@@

Under the conditions of the proposition, \llabel{pop_not_gaussian}
the parental traits, $Z^{[1]}_M$ and $Z^{[2]}_M$,
are also (jointly) Gaussian.
(This was not an assumption, however.)
If parents are chosen uniformly, then
the distribution of, say, $Z^{[1]}$ is just the population's trait distribution.
This seems wrong, since it is well-known that forces such as selection or migration
can make the population distribution non-Gaussian.
However, there is no contradiction:
we have stated the proposition only in terms of the \emph{first} generation
of offspring of unrelated parents \citep[as in][]{barton2017infinitesimal}.
Proposition~\ref{prop:parental_contribs} does not imply that
the distribution of traits in the offspring generation is Gaussian:
unless all offspring are unrelated,
it will be a collection of correlated Gaussians,
which is not itself Gaussian.
(As an extreme example, suppose that half the offspring are produced by the two parents
with largest trait values, and the other half by the parents with the smallest trait values;
then although each offspring's distribution about the midparent is Gaussian,
the offspring's population distribution is a mixture of two Gaussians
about these two midparents.)
What about further generations?
A similar statement to the proposition
holds for the offspring of two parents in an arbitrary pedigree,
using the Skitovich-Darmois theorem \citet{kagan1973characterization}
% except that the portion of the genome shared in common by the parents is unconstrained.
-- however, there is nothing to be gained by the additional generality.

@@

Although an infinitesimal model with non-Gaussian noise is incompatible with the usual additivity assumptions,
such models still exist (and might be useful);
next, we describe some of their properties.
As observed in \cite{turelli2017commentary},
there are at least three interpretations of what is called ``the infinitesimal model''. 
The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation,
assumes parental trait-independent Gaussian Mendelian sampling terms,
and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
As also discussed in \citet{turelli2017commentary}, \revpoint{1}{8}
this assumption is rather fragile --
a Gaussian distribution for the population only emerges under weak or Gaussian selection.
% However, the model has been used in influential work.
In principle, another factor that could produce a non-Gaussian population distribution
is if the Mendelian sampling terms are not Gaussian, as we consider here --
so, it is interesting to look for other situations where the population distribution is analytically tractable.

@@

First, consider the case without selection, so $\mu$ is a constant
which we may take equal to 1. 
(To remind the reader, our goal is therefore to describe \llabel{clarification1}
what forms the population distribution might take in a neutral model
if we assume only that the Mendelian sampling terms are independent of parental values,
and nothing else about the distribution of these terms or the initial population distribution.) 
Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta) ,
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) .
\]
Any fixed point $\phi(\theta)$ must be the limit of an arbitrary function, $\varphi(\theta)$, under repeated applications of $F$,
\ie \revpoint{2}{7}
\begin{equation}\label{eq:fpeq}
\begin{aligned}
	\phi(\theta) 
    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
\end{aligned}
\end{equation}
% Iterating $F$ $n$ times, we get
% \[
% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=0}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
% \]
% so that
% \[
% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
% \]
is a fixed point whenever this limit exists,
whereas any fixed point $\phi$ of $F$ satisfies  $\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$
for some function $\varphi(\theta)$.
As we are interested in identifying those fixed points that are characteristic functions, we henceforth assume, without loss of generality, that $\varphi$ is itself a characteristic function (if $\phi$ exists then we can always take $\varphi = \phi$) and that the limit $\lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$ exists.
